Converging High-Throughput and High-Performance Computing Paradigms: \\ A Case Study of Using Abstractions to Bridge the Divide

### 2017-03-17 Experiments.tex
Transition V to VI

In the batch queue,  queue time becomes dominant but, at the same time, we have more freedom to decide the parameters of the slot.

@@

In this section we describe experiments that have been performed by running the ATLAS workload on TITAN's batch queue by means of the next generation executor.   
The importance of these experiments is twofold: on the one side, we illustrate the performances of the NGE; on the other side, they help to characterize TITAN's batch queue and we show that it can be used as alternative to backfilling. Since we aim to study the response time of batch queue as it is perceived by a common user,  special privileges have been asked to system administrator.
%This is particularly important for the development of future execution strategies because queue time becomes dominant in

Experiments aim to show the number of completed events as function of the time. 
Each experiment corresponds to a number of different and independent runs that is meaningful from a statistical point of view. Each run consists in the sustained execution of the ATLAS workload for a pre-determined period of time \aanote{A week? 10 days?}. Since execution of events is carried on by means NGE jobs, each run is based on a script that maintains constant pressure on TITAN's batch queue.
The main tasks of the script are to:  i) guarantee that there are constantly four NGE jobs eligible for running on TITAN's queue (note that four jobs is the maximum allowed in TITAN's batch queue \ref{}); ii)  bind Athena-MP jobs to the NGE jobs; iii) perfom garbage collection. No limit in the number of NGE jobs is imposed.

The simulation of events is due to the following execution model: the script binds a number of Athena-MP jobs that is equal to the number of nodes on which NGE job is running; then, NGE job spawns the Athena-MP jobs on the nodes; finally, the Athena-MP jobs will simulate a number of events that is proportional to the wall-time of the NGE job. As a consequence, the number of nodes determines the number of events that can be simulated concurrently whereas the wall-time determines the number of events that can be simulated one after the other. These two parameters together determine the size of the NGE job.

The size of the NGE job is constant within an experiment. Thus, each experiment is fully characterized by the size of the NGE jobs that it send on the nodes.

Each execution of Athena-MP is guaranteed to be successfull and, on average, comparable with those executed by PanDA in production.

Table \ref{tab:experiments} provides the list of the experiments that have been performed.

\begin{table*}
\begin{center}
\begin{tabular}{|c|l|l|l|l|}
  \hline
Experiment &\# nodes &wall-time (hours) & run duration(days) & \# runs   \\
\hline
1 & &  & &   \\
2 & &  & &  \\
3 & &  & &  \\
4 & &  & &  \\
\hline
\end{tabular}
\end{center}
\caption{List of experiments}\label{tab:experiments}
\end{table*}

Each experiment has been designed by choosing the maximum number of nodes and the maximum wall-time of each TITAN bin \ref{} \aanote{Are we really doing this???}. The length of the experiment has been decided.....\aanote{It was 10 times the wall-time or something like that}.

Figure \ref{fig:avtEvts} \aanote{(DRAFT)}depicts average number of completed events as function of the time. Figure \ref{fig:avtTimes} \aanote{(DRAFT)} depicts the average queue time and execution time per experiment.

\begin{figure}[!htb]
        \includegraphics[width=0.48\textwidth]{./figures/draft/mean.pdf}
    \caption{Average number of completed as function of the time.}
\label{fig:avtEvts}
\end{figure}

\begin{figure}[!htb]
        \includegraphics[width=0.48\textwidth]{./figures/draft/AvgT.pdf}
    \caption{Average queue time and execution time of the experiments.}
\label{fig:avtTimes}
\end{figure}

Because of the definition of the experiment the results shown in Figure \ref{} are alike (but not entirely comparable to) those provided by in Figure \ref{} where the performances of backfilling are shown.

>>>>

% In the batch queue,  queue time becomes dominant but, at the same time, we
% have more freedom to decide the parameters of the slot.

@@

In this section we illustrate the performances of the NGE by means of experiments.
Experiments aim to determine the performances of the NGE and to show how it provides a negligible overhead while introducing new functionalities.

All the experiments have been performed on TITAN's batch queue. This allowed us to decide the pilot size in terms of number of nodes and wall-time. For the same reason, the completion time of each run of our experiments is dominated by the queue time. However, since we are interested only in the performances of the NGE, we present only data related to the execution time of the pilot and the tasks that are run within it.

We test the NGE pilot for weak scalability, weak scalability with resubmission, strong scalability and we provide a proof of concept about how the NGE is able to run heterogeous workloads.

\subsubsection{Weak scalability}
\aanote{I am using the term simulation....I will change it once we decide how to call the CU}
The experiment consists in running as many tasks as the number of nodes controlled by the pilot.
Tasks consists in Athena-MP job that simulate 100 events taken from ATLAS workload.
Each Athena-MP job uses 16 cores which coincide with the number of cores available on TITAN's nodes.
In order to minimize measurement noises, we used the very same simulation for all the tasks. Therefore, ideally all the tasks should have the same execution time which corresponds to ~$80$ minutes.

Tasks do not experience queue within the pilot since there is one node for each ATLAS simulation. Therefore, delays are of only three factors: i) the bootstrapping of the pilot on the nodes; ii) the manager that has to dispatch tasks to the pilot; iii) the time that the agent requires to spawn all of simulations on the nodes.

We tested different pilot sizes, i.e. : 250, 500, 1000 and 2000. For all of them the walltime was 2 hours.

Figure \ref{fig:weakScal1} depicts the average pilot duration, the average execution time of each tasks and the average time required to send a tasks for execution.

\begin{figure}[!htb]
        \includegraphics[width=0.5\textwidth]{./figures/NGE/Weak.pdf}
    \caption{Average pilot execution time, average tasks execution time and average launching time for pilot sizes: 250, 500, 1000 and 2000.}
\label{fig:weakScal1}
\end{figure}
\aanote{TO be commented as soon as the figure becomes definitive. Note that It would be better if we show the ratio between pilot duration and execution time}
\subsubsection{Weak scalability with resubmission }
This experiment is similar to the one presented above but in this case we want to test also the impact of submitting new simulations in place of those that end. Thus, we send to the pilot a number of tasks that is five times larger than the number of nodes. The reason behind this experiment is to stress the manager and the agent by pushing new tasks while others are ending their execution in such a way to increase the concurrency of these two components.

In order to perform the experiments in a reasonable amount of time, we reduced  the number of events that are simulated by Athena-MP to sixteen. This allows us to complete a task in ~25 minutes only. Note that the number of events has been chosen in such a way that all the cores of a node run one event.

We tested different pilot sizes, i.e. : 250, 500, 1000 and 2000. For all of them the walltime was 3 hours.

Figure \ref{fig:weakScal2} \aanote{(DRAFT)}  depicts the average pilot duration, the average execution time of each tasks and the average time required to send a tasks for execution.

\begin{figure}[!htb]
        \includegraphics[width=0.48\textwidth]{./figures/draft/AvgT.pdf}
    \caption{Average pilot execution time, average tasks execution time and average launching time for pilot sizes: 250, 500, 1000 and 2000.}
\label{fig:weakScal2}
\end{figure}

\subsubsection{Strong scability}

\subsubsection{Heterogeneous execution}

### 2017-03-17 introduction.tex
Clean up

\section{Introduction}\label{sec:intro}

The Large Hadron Collider (LHC) is created to explore the fundamental properties
of matter for the next decades. Since LHC start-up in 2009, the experiments
collected and distributed hundreds of petabytes of data worldwide to hundreds of
computer centers. Thousands of physicists analyze petascale data volumes daily.
One of the LHC experiments, the ATLAS~\cite{Aad:2008}, utilizes the Production
and Distributed Analysis (PanDA) workload management system~\cite{Maeno2011}
(WMS) for distributed data processing and analysis. The ATLAS Computing
model~\cite{Atlas2005} is based on a Grid paradigm~\cite{Foster:1998}, with
multilevel, hierarchically distributed computing and storage resources. PanDA
has been developed to meet growing ATLAS production and analysis requirements
for a data-driven workload management system capable of operating at LHC data
processing scale.

PanDA has a highly scalable architecture. Scalability has been demonstrated in
ATLAS through the rapid increase in usage over the past several years of
operations, and PanDA is expected to meet the continuously growing computing
requirements of ATLAS over the next decade. PanDA was designed to have the
flexibility to adapt to emerging computing technologies in processing, storage,
networking as well as the underlying software stack. This flexibility has also
been successfully demonstrated through the past six years: computing centers in
ATLAS, spanning many continents, were seamlessly integrated into PanDA\@.

PanDA manages a wide spectrum of workloads, ranging from raw data processing to
Monte Carlo simulation and user analysis, while constantly evolving to meet
rapidly changing science needs. Today, PanDA serves several thousand users,
managing job distribution to hundreds of ATLAS sites with more than 100,000 CPU
cores which process more than a million jobs per day (Figure~\ref{fig:daily}).

>>>>

The Large Hadron Collider (LHC) was created to explore the fundamental
properties of matter for the next decades.  Since LHC start-up in 2009,
multiple experiments  at LHC have collected and distributed hundreds of
petabytes of data worldwide to hundreds of computer centers. 
Thousands of physicists analyze petascale data volumes daily.
One of the LHC experiments, the ATLAS~\cite{Aad:2008}, utilizes the Production
and Distributed Analysis (PanDA) workload management system~\cite{Maeno2011}
(WMS) for distributed data processing and analysis. The ATLAS Computing
model~\cite{Atlas2005} is based on a Grid paradigm~\cite{Foster:1998}, with
multilevel, hierarchically distributed computing and storage resources. PanDA
has been developed to meet growing ATLAS production and analysis requirements
for a data-driven workload management system capable of operating at LHC data
processing scale.

% PanDA has a highly scalable architecture. Scalability has been demonstrated in
% ATLAS through the rapid increase in usage over the past several years of
% operations, and PanDA is expected to meet the continuously growing computing
% requirements of ATLAS over the next decade. PanDA was designed to have the
% flexibility to adapt to emerging computing technologies in processing, storage,
% networking as well as the underlying software stack. This flexibility has also
% been successfully demonstrated through the past six years: computing centers in
% ATLAS, spanning many continents, were seamlessly integrated into PanDA\@.

% PanDA manages a wide spectrum of workloads, ranging from raw data processing to
% Monte Carlo simulation and user analysis, while constantly evolving to meet
% rapidly changing science needs.

Today, PanDA serves several thousand users, managing job distribution to
hundreds of ATLAS sites with more than 100,000 CPU cores which process more
than a million jobs per day (Figure~\ref{fig:daily}).
